---
title: "Reported Domestic Violence in Chicago City, IL"
author: Ting Chang
subtitle: "Before and After the COVID-19 Lockdown"
---

# Introduction

The surging cases of COVID-19 in spring 2020 has caused governments around the world to announce the lockdown policy (stay home order) in order to slow down the spread of pandemic. While COVID-19 is a worldwide health crisis, some researchers have expressed concern for another crucial health threat, domestic violence, which is expected to increase while more people were constrained at home.

This project aims to explore the spatial and temporal distribution of the domestic violence in Chicago City, Illinois using the crime reported data. In particularly, to examine whether the reported domestic violence has increased as expected after the implementation of lockdown.


# Materials and methods
## Set up
In order to run the code below, firstly, we need to load the following packages in R, you might need to install some of them in ahead. If your environment and locale is not set to English, utilizing the two lines start with `Sys.`, it would help plotting the date label in English. Otherwise, please comment out these two lines.

```{r setup, message=F, warning=F, results='hide'}
library(tidyverse)
library(tidycensus)
library(kableExtra)
library(ggplot2)
library(sf)
library(RColorBrewer)
library(viridis)
library(gridExtra)
library(classInt)

knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling

# For plotting date label in English
Sys.setenv("LANGUAGE"="En")
Sys.setlocale("LC_ALL", "English")

# Changing environment and locale language learned from: 
# https://stackoverflow.com/questions/15438429/axis-labels-are-not-plotted-in-english
```


## Data Import and Processing
### Chicago City's Census Block Group Shapefile
The Chicago City's census block group shapefile is provided in the data folder. I have obtained it previously from the census TIGER data.

```{r shapefile, message=F, warning=F, results='hide'}
# Load chicago CBG shapefile in R
chicago_cbg <- st_read("data/chicago_cbg.shp") %>%
  select(CensusBloc)
```


### Census Data
We will download several datasets from the American Community Survey 5-Year data at the census block group level using the lovely package, tidycensus. Since the newest released estimated data is the 2018 year version, we will download data from 2018.

```{r census, message=F, warning=F, results='hide'}
# For searching variables from ACS-5yr 2018
v18 <- load_variables(2018, "acs5", cache = TRUE)

# Downloading population data
population_18 <- get_acs(geography = "block group",
                         variables = c(population = "B01003_001"),
                         state = "IL",
                         county = c("Cook", "DuPage"),
                         year = 2018)

# Downloading median household income data
median_income_18 <- get_acs(geography = "block group",
                            variables = c(medincome = "B19013_001"),
                            state = "IL",
                            county = c("Cook", "DuPage"),
                            year = 2018)

# Downloading employment data
employment_18 <- get_acs(geography = "block group",
                         variables = c(total_in_labor = "B23025_002", unemployed = "B23025_005"),
                         state = "IL",
                         county = c("Cook", "DuPage"),
                         year = 2018,
                         output = "wide")
```

Afterwards, we will crop the obtained census data to the Chicago City using the pre-loaded shapefile and slightly tidy up our data.

```{r census-cropped, message=F, warning=F, results='hide'}
# Population data
population_18 <- left_join(chicago_cbg, population_18, by = c("CensusBloc" = "GEOID")) %>%
  select(CensusBloc, population = estimate)

# Median household income data
median_income_18 <- left_join(chicago_cbg, median_income_18, by = c("CensusBloc" = "GEOID")) %>%
  select(CensusBloc, median_income = estimate)

# Unemployment rate data
employment_18 <- left_join(chicago_cbg, employment_18, by = c("CensusBloc" = "GEOID")) %>%
  # calculate the unemployment rate by dividing the unemployed population by the total in labor population
  select(CensusBloc, total_in_labor = total_in_laborE, unemployed = unemployedE) %>%
  mutate(unemployment_rate = unemployed/total_in_labor)
```


### Crime Reported Data {.tabset}
The crime reported data will be obtained from the [Chicago Data Portal](https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-Present/ijzp-q8t2) via Socrata Open Data API.

```{r crime-report, message=F, warning=F, results='hide'}
# Download crime reported data and load it in R
dataurl = "https://data.cityofchicago.org/resource/ijzp-q8t2.csv?$order=Date DESC&$limit=460000&$offset=20000"
tdir = tempdir()
download.file(dataurl, destfile = file.path(tdir, "chicago_crime.csv"))

chicago_crime <- read_csv(paste(tdir,"/chicago_crime.csv", sep = ""))

# Tidy the data
chicago_crime_clean <- chicago_crime %>%
  mutate(date = as.Date(date, "%Y.%m.%d")) %>%
  filter(as.Date(date) >= "2018-12-31" & as.Date(date) <= "2020-10-04") %>%
  select(-block, -iucr, -beat, -district, -ward, -community_area, -x_coordinate, -y_coordinate, -year, -location) %>%
  drop_na(latitude, longitude)

# The SODA API docs:
# https://dev.socrata.com/foundry/data.cityofchicago.org/ijzp-q8t2
# The SODA API query docs:
# https://dev.socrata.com/docs/queries/
# Learn paste() function from:
# https://stackoverflow.com/questions/27378116/how-to-change-a-file-path-in-r-with-a-constant
```
Then, we will apply the spatial join function `st_join` to compute which census block group did each crime incident occur.

```{r crime-cbg, message=F, warning=F, results='hide'}
# Convert crime reported data to sf object
chicago_crime_clean <- st_as_sf(chicago_crime_clean, coords = c("longitude", "latitude"),
                                crs = st_crs(chicago_cbg))

# Spatial joint the crime reported data with census block groups shapefile
chicago_crime_cbg <- st_join(chicago_crime_clean, chicago_cbg) %>%
  st_set_geometry(NULL)
```

All the pre-processing works for crime reported data have been done now. Next, we will work on deriving the crime rate. Furthermore, we will extract the domestic violence reported data, calculate the domestic violence counts by week, as well as derive the domestic violence rate.


#### Crime Rate
To obtain the crime rate for each census block group, we will first compute the number of crime that had occurred within each census block group in different months. Then we will normalize the crime counts by the population size of each census block group.

```{r crime-rate, message=F, warning=F, results='hide'}
cbg_crime_march19 <- chicago_crime_cbg %>%
  # filter the data by month
  filter(date >= "2019-03-01" & date <= "2019-03-31") %>%   
  # count the number of crimes occurred in each CBG
  group_by(CensusBloc) %>%
  tally(name = "crime_count") %>%
  # join crime data with population dataframe
  right_join(population_18, by = "CensusBloc") %>%
  # normalize crime by population size of each CBG
  mutate(crime_count = replace_na(crime_count, 0), crime_rate = crime_count/population) %>%
  # convert dataframe to sf object
  st_as_sf(crs=st_crs(chicago_cbg))

cbg_crime_april19 <- chicago_crime_cbg %>%
  filter(date >= "2019-04-01" & date <= "2019-04-30") %>%   
  group_by(CensusBloc) %>%
  tally(name = "crime_count") %>%
  right_join(population_18, by = "CensusBloc") %>%
  mutate(crime_count = replace_na(crime_count, 0), crime_rate = crime_count/population) %>%
  st_as_sf(crs=st_crs(chicago_cbg))

cbg_crime_march20 <- chicago_crime_cbg %>%
  filter(date >= "2020-03-01" & date <= "2020-03-31") %>%   
  group_by(CensusBloc) %>%
  tally(name = "crime_count") %>%
  right_join(population_18, by = "CensusBloc") %>%
  mutate(crime_count = replace_na(crime_count, 0), crime_rate = crime_count/population) %>%
  st_as_sf(crs=st_crs(chicago_cbg))

cbg_crime_april20 <- chicago_crime_cbg %>%
  filter(date >= "2020-04-01" & date <= "2020-04-30") %>%   
  group_by(CensusBloc) %>%
  tally(name = "crime_count") %>%
  right_join(population_18, by = "CensusBloc") %>%
  mutate(crime_count = replace_na(crime_count, 0), crime_rate = crime_count/population) %>%
  st_as_sf(crs=st_crs(chicago_cbg))

# replace_na learned from the official document:
# https://dplyr.tidyverse.org/reference/tally.html
# Count observations learned from official document:
# https://dplyr.tidyverse.org/reference/tally.html
```


#### Domestic Violence Reported Data
To obtain the domestic violence reported data, we will utilize the domestic violence column, location description column, as well as primary crime type column to filter the domestic violence reported data.

```{r dv-process, message=F, warning=F, results='hide'}
# Filter out only the DV occurred in residential place
list_location <- c("APARTMENT",
                   "CHA APARTMENT",
                   "CHA HALLWAY / STAIRWELL / ELEVATOR",
                   "CHA HALLWAY/STAIRWELL/ELEVATOR",
                   "CHA PARKING LOT",
                   "CHA PARKING LOT / GROUNDS",
                   "CHA PARKING LOT/GROUNDS",
                   "COACH HOUSE",
                   "DRIVEWAY - RESIDENTIAL",
                   "HOUSE",
                   "NURSING / RETIREMENT HOME",
                   "NURSING HOME/RETIREMENT HOME",
                   "RESIDENCE",
                   "RESIDENCE - GARAGE",
                   "RESIDENCE - PORCH / HALLWAY",
                   "RESIDENCE - YARD (FRONT / BACK)",
                   "RESIDENCE PORCH/HALLWAY",
                   "RESIDENCE-GARAGE",
                   "RESIDENTIAL YARD (FRONT/BACK)")

# Filter the crime type
list_crime_type <- c("ARSON",
                     "ASSAULT",
                     "BATTERY",
                     "BURGLARY",
                     "CRIM SEXUAL ASSAULT",
                     "CRIMINAL SEXUAL ASSAULT",
                     "DOMESTIC VIOLENCE",
                     "HOMICIDE",
                     "INTIMIDATION",
                     "KIDNAPPING",
                     "OBSCENITY",
                     "OFFENSE INVOLVING CHILDREN",
                     "OTHER OFFENSE",
                     "SEX OFFENSE",
                     "STALKING")

chicago_dv_cbg <- chicago_crime_cbg %>%
  filter(domestic == TRUE & location_description %in% list_location & primary_type %in% list_crime_type)
```


#### Domestic Violence Count by Week
The domestic violence count aggregated by week can be easily obtained by means of the `dplyr` functions.

```{r dv-count-week, message=F, warning=F, results='hide'}
dv_count_week <- chicago_dv_cbg %>%
  group_by(week = cut(date, "week")) %>%
  tally(name = "dv_count") %>%
  mutate(week = as.Date(week))

# Aggregate by week learned from:
# https://stackoverflow.com/questions/40554231/dplyr-lubridate-how-to-aggregate-a-dataframe-by-week/40554522
```


#### Domestic Violence Rate
The domestic violence rate will be derived using the same method as crime rate. 

```{r dv-rate, message=F, warning=F, results='hide'}
cbg_dv_march19 <- chicago_dv_cbg %>%
  filter(date >= "2019-03-01" & date <= "2019-03-31") %>%   
  group_by(CensusBloc) %>%
  tally(name = "dv_count") %>%
  right_join(population_18, by = "CensusBloc") %>%
  mutate(dv_count = replace_na(dv_count, 0), dv_rate = dv_count/population) %>%
  st_as_sf(crs=st_crs(chicago_cbg))

cbg_dv_april19 <- chicago_dv_cbg %>%
  filter(date >= "2019-04-01" & date <= "2019-04-30") %>%
  group_by(CensusBloc) %>%
  tally(name = "dv_count") %>%
  right_join(population_18, by = "CensusBloc") %>%
  mutate(dv_count = replace_na(dv_count, 0), dv_rate = dv_count/population) %>%
  st_as_sf(crs=st_crs(chicago_cbg))

cbg_dv_march20 <- chicago_dv_cbg %>%
  filter(date >= "2020-03-01" & date <= "2020-03-31") %>%
  group_by(CensusBloc) %>%
  tally(name = "dv_count") %>%
  right_join(population_18, by = "CensusBloc") %>%
  mutate(dv_count = replace_na(dv_count, 0), dv_rate = dv_count/population) %>%
  st_as_sf(crs=st_crs(chicago_cbg))

cbg_dv_april20 <- chicago_dv_cbg %>%
  filter(date >= "2020-04-01" & date <= "2020-04-30") %>%
  group_by(CensusBloc) %>%
  tally(name = "dv_count") %>%
  right_join(population_18, by = "CensusBloc") %>%
  mutate(dv_count = replace_na(dv_count, 0), dv_rate = dv_count/population) %>%
  st_as_sf(crs=st_crs(chicago_cbg))
```

```{r classint, eval=F, echo=F}
# # ClassInt package could convert continuous variable to categorical
# 
# # Having issue: when using "jenks" style, the brackets are forced to be right-closed, if need to include the minimum value, in my case zero, need to use, e.g., min(variable)-0.001.
# 
# # Also need to research on how to change the legend, the current method to plot out choropleth map, the legend is pretty messy.
# 
# dv_breaks <- classIntervals(c(0,cbg_dv_march20$dv_rate), n=6, style = "jenks")
# dv_breaks <- classIntervals(cbg_dv_march20$dv_rate, n=6, style = "fixed",
#                             fixedBreaks=c(0,0.000555,0.001862,0.003807,0.007212,0.015038,0.025271))
# 
# cbg_dv_march20 <- mutate(cbg_dv_march20, dv_rate_cat = cut(dv_rate, dv_breaks$brks))
# 
# # https://cengel.github.io/R-spatial/mapping.html
```

# Results

[~200 words]

Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.


Lineplot of domestic violence count

```{r lineplot, message=F, warning=F, fig.align='center', fig.width=10, fig.height=5}
dv_col <- brewer.pal(8, "Dark2")[2]

ggplot(data = dv_count_week, aes(x = week, y = dv_count, group = 1)) +
  geom_line(color = dv_col, lwd = 0.8, alpha=0.7) +
  geom_point(color = dv_col, alpha=0.7) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 months") +
  labs(title="Reported Domestic Violence Over Time", x = "Month", y = "Count of Domestic Violence") +
  theme_light() +
  theme(axis.text.x = element_text(angle=45, hjust = 1))
```

```{r, fig.show='hold', fig.align='center', fig.width=12, fig.height=15, dpi=500}
map_dv_march19 <- ggplot(data=cbg_dv_march19) +
  geom_sf(aes(fill = dv_rate), lwd=0.3) +
  scale_fill_viridis(option="magma", direction = -1) +
  theme_minimal() +
  theme(legend.position = 'bottom') +
  labs(title = "Domestic violence rate in March 2019",
       fill = "Domestic violence rate")

map_dv_april19 <- ggplot(data=cbg_dv_april19) +
  geom_sf(aes(fill = dv_rate), lwd=0.3) +
  scale_fill_viridis(option="magma", direction = -1) +
  theme_minimal() +
  theme(legend.position = 'bottom') +
  labs(title = "Domestic violence rate in April 2019",
       fill = "Domestic violence rate")

map_dv_march20 <- ggplot(data=cbg_dv_march20) +
  geom_sf(aes(fill = dv_rate), lwd=0.3) +
  scale_fill_viridis(option="magma", direction = -1) +
  theme_minimal() +
  theme(legend.position = 'bottom') +
  labs(title = "Domestic violence rate in March 2020",
       fill = "Domestic violence rate")

map_dv_april20 <- ggplot(data=cbg_dv_april20) +
  geom_sf(aes(fill = dv_rate), lwd=0.3) +
  scale_fill_viridis(option="magma", direction = -1) +
  theme_minimal() +
  theme(legend.position = 'bottom') +
  labs(title = "Domestic violence rate in April 2020",
       fill = "Domestic violence rate")

grid.arrange(map_dv_march19, map_dv_march20, map_dv_april19, map_dv_april20, nrow=2, ncol=2)

# Reverse color palette:
# https://stackoverflow.com/questions/45868625/how-to-reverse-the-default-color-palette-for-ggplot2
# grid.arrange refers to:
# https://cran.r-project.org/web/packages/egg/vignettes/Ecosystem.html
```

```{r, message=F, warning=F, fig.align='center', fig.width=12, fig.height=6}
median_col <- brewer.pal(8, "Pastel2")[1]

hist_median18 <- ggplot(data = median_income_18, aes(x=median_income)) +
  geom_histogram(fill=median_col, col="grey50", binwidth = 10000) +
  theme_classic() +
  labs(title = "Histogram of Median Household Income", x="Median Household Income", y="Count")

map_median18 <- ggplot(data=median_income_18) +
  geom_sf(aes(fill = median_income), lwd=0.3) +
  scale_fill_viridis(option="viridis", direction = -1) +
  theme_minimal() +
  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank()) +
  labs(title = "Median Household Income in 2018",
       fill = "Median household\nincome")
# make sure the NA CBGs

grid.arrange(map_median18, hist_median18, ncol=2, widths=2:1)
```

```{r, message=F, warning=F, fig.align='center', fig.width=12, fig.height=6}
unemployed_col <- brewer.pal(8, "Pastel2")[2]

hist_unemployed18 <- ggplot(data = employment_18, aes(x=unemployment_rate)) +
  geom_histogram(fill=unemployed_col, col="grey50", binwidth = 0.04) +
  theme_classic() +
  labs(title = "Histogram of Unemployment Rate", x="Unemployment Rate", y="Count")

map_unemployed18 <- ggplot(data=employment_18) +
  geom_sf(aes(fill = unemployment_rate), lwd=0.3) +
  scale_fill_viridis(option="viridis", direction = -1) +
  theme_minimal() +
  theme(axis.line = element_blank(), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank()) +
  labs(title = "Unemployment Rate in 2018",
       fill = "Unemployment rate")

grid.arrange(map_unemployed18, hist_unemployed18, ncol=2, widths=2:1)
```


# Conclusions

[~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

All sources are cited in a consistent manner
